---
title: "222_Project"
author: "Colby Wilkinson"
date: "11/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
set.seed(35)
```

#Bayesian estimator and confidence interval code (DAN)

```{r}
# n is sample size from exponential distribution
# replicates is number of samples from exponential distribution
# rate is exponential rate parameter

simulate_exponential_gamma <- function(n, rate, replicates){
  
  alpha <- c()
  beta <- c()
  for(i in 1:replicates){ 
    x <- rexp(n, rate)
    beta_i <- sum(x)
    alpha_i <- n + 1
    alpha <- c(alpha, alpha_i)
    beta <- c(beta, beta_i)
  }
  
  
  hits <- c()
  for(i in 1:replicates){
    GAMMAS <- qgamma(c(0.025,0.975), alpha[i], beta[i])
    if(rate > GAMMAS[2] || rate < GAMMAS[1]){
      hits = c(hits,0) 
    } else {
      hits = c(hits,1)
    }
  }
  
  bias <- alpha/beta - rate
  variance <- alpha/beta^2
  MSE <- variance + bias^2
  
  return(list(coverage = mean(hits), bias = bias, variance = variance, MSE = MSE))  
  
}  


# testing
simulations <- simulate_exponential_gamma(n = 10000, rate = 1, replicates = 1000)
mean(simulations$coverage)
mean(simulations$MSE)
mean(simulations$bias)

# coverages and measurements over sample size
n <- 5:100
data <- lapply(n, simulate_exponential_gamma, rate = 5, replicates = 1000)
coverages <- c()
mean_bias <- c()
mean_variance <- c()
mean_MSE <- c()
for(i in 1:length(data)){
  coverages <- c(coverages, data[[i]]$coverage)
  mean_bias <- c(mean_bias, mean(data[[i]]$bias))
  mean_variance <- c(mean_variance, mean(data[[i]]$variance))
  mean_MSE <- c(mean_MSE, median(data[[i]]$MSE))
}


manyN <- data.frame(n, coverages, mean_bias, mean_variance, mean_MSE)
#coverages over sample size
manyN %>% ggplot(aes(x = n, y = coverages)) + geom_point(alpha = 0.5) + geom_smooth(method = 'loess')
#bias, mse, variance over sample size
manyN %>% ggplot() + geom_smooth(aes(n, mean_bias), col = 'red') + 
  geom_smooth(aes(n,mean_variance), col = 'blue') + 
  geom_smooth(aes(n,mean_MSE), col = 'green')


# changing rates fixed sample size
rate <- seq(0.01, 20, length.out = 100)
data2 <- lapply(rate, simulate_exponential_gamma, n = 50, replicates = 1000)

coverages <- c()
mean_bias <- c()
mean_variance <- c()
mean_MSE <- c()
for(i in 1:length(data2)){
  coverages <- c(coverages, data2[[i]]$coverage)
  mean_bias <- c(mean_bias, mean(data2[[i]]$bias))
  mean_variance <- c(mean_variance, mean(data2[[i]]$variance))
  mean_MSE <- c(mean_MSE, median(data2[[i]]$MSE))
}

manyR10 <- data.frame(rate, coverages, mean_bias, mean_variance, mean_MSE)
#coverage over rates
manyR10 %>% ggplot(aes(x = rate, y = coverages)) + geom_smooth(method = 'loess')
#bias, mse, variance over rates
manyR10 %>% ggplot() + geom_smooth(aes(rate, mean_bias), col = 'red') + 
  geom_smooth(aes(rate, mean_variance), col = 'blue') + 
  geom_smooth(aes(rate, mean_MSE), col = 'green')


data3 <- lapply(rate, simulate_exponential_gamma, n = 50, replicates = 1000)

coverages <- c()
mean_bias <- c()
mean_variance <- c()
mean_MSE <- c()
for(i in 1:length(data2)){
  coverages <- c(coverages, data3[[i]]$coverage)
  mean_bias <- c(mean_bias, mean(data3[[i]]$bias))
  mean_variance <- c(mean_variance, mean(data3[[i]]$variance))
  mean_MSE <- c(mean_MSE, median(data3[[i]]$MSE))
}

manyR50 <- data.frame(rate, coverages, mean_bias, mean_variance, mean_MSE)
manyR50 %>% ggplot(aes(x = rate, y = coverages)) + geom_point(alpha = 0.5) + geom_smooth(method = 'loess')
manyR50 %>% ggplot() + geom_smooth(aes(rate, mean_bias), col = 'red') + 
  geom_smooth(aes(rate, mean_variance), col = 'blue') + 
  geom_smooth(aes(rate, mean_MSE), col = 'green')


data4 <- lapply(rate, simulate_exponential_gamma, n = 50, replicates = 1000)

coverages <- c()
mean_bias <- c()
mean_variance <- c()
mean_MSE <- c()
for(i in 1:length(data2)){
  coverages <- c(coverages, data4[[i]]$coverage)
  mean_bias <- c(mean_bias, mean(data4[[i]]$bias))
  mean_variance <- c(mean_variance, mean(data4[[i]]$variance))
  mean_MSE <- c(mean_MSE, median(data4[[i]]$MSE))
}

manyR25 <- data.frame(rate, coverages, mean_bias, mean_variance, mean_MSE)
manyR25 %>% ggplot(aes(x = rate, y = coverages)) + geom_point(alpha = 0.5) + geom_smooth(method = 'loess')
manyR25 %>% ggplot() + geom_smooth(aes(rate, mean_bias), col = 'red') + 
  geom_smooth(aes(rate, mean_variance), col = 'blue') + 
  geom_smooth(aes(rate, mean_MSE), col = 'green')


```


# delta method, fixed n-size coverage probabilities for parameters 0.1 to 10

```{r}

#Delta function for coverage probabilities

delta_ci <- function(observations, parameter, n_size){


dat <- data.frame(observation = seq(1,observations,1), true_parameter = parameter, n_size = rep(n_size,observations))


for (i in 1:length(dat$observation)){
  dat$estimate[i] <- 1/mean(rexp(n = dat$n_size[i], rate = dat$true_parameter[i]))
  
  dat$lower[i] <- dat$estimate[i] - (1.96*dat$estimate[i])/sqrt(dat$n_size[i])
  
  dat$upper[i] <- dat$estimate[i] + (1.96*dat$estimate[i])/sqrt(dat$n_size[i])
  
  dat$hit[i] <- ifelse(dat$true_parameter[i] > dat$lower[i] & dat$true_parameter[i] < dat$upper[i], 1, 0)
} 
return(sum(dat$hit)/observations)
}



# Delta Simulation

dat_delta_parameter <- data.frame(parameter = seq(0.01, 20, length.out = 100))

for (i in 1:length(dat_delta_parameter$parameter)) {

  dat_delta_parameter$cov_prob[i] <- delta_ci(1000, dat_delta_parameter$parameter[i],50)
}


#Delta plot vs. paramter (n = 100)

dat_delta_parameter %>%
  ggplot(aes(x = parameter, y = cov_prob)) +
  geom_point()

```




# exact confidence interval *

## 1/mean as estimator, exact confidence intervals, fixed n-size coverage probabilities for parameters 0.1 to 10
 
```{r}

#exact ci function

exact_ci <- function(observations, parameter, n_size){


dat <- data.frame(observation = seq(1,observations,1), true_parameter = parameter, n_size = rep(n_size, observations))


for (i in 1:length(dat$observation)){
  
  dat$estimate[i] <- 1/mean(rexp(n = dat$n_size[i], rate = dat$true_parameter[i]))
  
  dat$lower[i] <- qchisq(0.025, 2*dat$n_size[i])/(2*dat$n_size[i]*(1/dat$estimate[i]))
  
  dat$upper[i] <- qchisq(0.975, 2*dat$n_size[i])/(2*dat$n_size[i]*(1/dat$estimate[i]))
  
  dat$hit[i] <- ifelse(dat$true_parameter[i] > dat$lower[i] & dat$true_parameter[i] < dat$upper[i], 1, 0)
} 
return(sum(dat$hit)/observations)
}


#exact ci simulation for 
 
dat_exact_parameter <- data.frame(parameter = seq(0.1, 20, length.out = 100))

for (i in 1:length(dat_exact_paramter$parameter)) {

  dat_exact_parameter$cov_prob[i] <- exact_ci(1000, dat_exact_paramter$parameter[i], 50)
}

dat_exact_parameter %>%
  ggplot() +
  geom_point(aes(x = parameter, y = cov_prob))

```


# delta ci, variable n-size coverage probabilities for parameters = 5

```{r}


dat_delta_n <- data.frame(n_size = seq(5, 100, 1))

for (i in 1:length(dat_delta_n$n_size)) {

  dat_delta_n$cov_prob[i] <- delta_ci(1000, 5, dat_delta_n$n_size[i])
}

```

# exact ci as estimator, variable n-size coverage probabilities for parameter = 5

```{r}

dat_exact_n <- data.frame(n_size = seq(5,100, 1))

for (i in 1:length(dat_exact_n$n_size)) {

  dat_exact_n$cov_prob[i] <- exact_ci(1000, 5, dat_exact_n$n_size[i])
}

```



#CI methods plotted together

```{r}


ggplot() +
  geom_smooth(data = manyR50, aes(x = rate, y = coverages), col = "yellow", se = FALSE) +
  geom_smooth(data = dat_delta_parameter, aes(x = parameter, y = cov_prob), col = "red", se = FALSE) +
  geom_smooth(data = dat_exact_parameter, aes(x = parameter, y = cov_prob), col = "green", se = FALSE) +
  scale_y_continuous(limits = c(0.9, 1), breaks = seq(0.9, 1, 0.01)) +
  xlab("Theta") +
  ylab("Coverage Probability")

ggplot() +
  geom_smooth(data = manyN, aes(x = n, y = coverages), col = "yellow", se = FALSE) +
  geom_smooth(data = dat_delta_n, aes(x = n_size, y = cov_prob), col = "red", se = FALSE) +
  geom_smooth(data = dat_exact_n, aes(x = n_size, y = cov_prob), col = "green", se = FALSE) +
  scale_y_continuous(limits = c(0.9, 1), breaks = seq(0.9, 1, 0.01)) +
  scale_x_continuous(limits = c(5, 100), breaks = seq(10,100,10)) +
  xlab("n-size") +
  ylab("Coverage Probability")



```




#Parameters


```{r}

estimators <- function(observations, parameter, n_size){

dat <- data.frame(observation = seq(1,observations,1), true_parameter = rep(parameter, observations), n_size = rep(n_size,observations))

for (i in 1:length(dat$observation)){
  
  sample_exp <- rexp(n = dat$n_size[i], rate = dat$true_parameter[i])
  
  dat$mle[i] <- 1/mean(sample_exp)
  dat$med[i] <- log(2)/median(sample_exp)
  
  dat$mle_bias[i] <- dat$mle[i] - dat$true_parameter[i]
  dat$med_bias[i] <- dat$med[i] - dat$true_parameter[i]
  
  dat$mle_sq_error[i] <- (dat$true_parameter[i] - dat$mle[i])^2
  dat$med_sq_error[i] <- (dat$true_parameter[i] - dat$med[i])^2
  } 



mle_mse <- mean(dat$mle_sq_error)
med_mse <- mean(dat$med_sq_error)

mle_var <- var(dat$mle)
med_var <- var(dat$med)

mle_avg_bias <- mean(dat$mle_bias)
med_avg_bias <- mean(dat$med_bias)

return(data.frame(mle_mse = mle_mse, med_mse = med_mse, mle_var = mle_var, med_var = med_var, mle_avg_bias = mle_avg_bias, med_avg_bias = med_avg_bias))

}


```

# Estimates mse, bias, variance over parameters

```{r}

dat_parameter <- data.frame(parameter = seq(0.01, 20, length.out = 100))

dat_sim_parameter <- data.frame(mle_mse = NULL, med_mse = NULL, mle_var = NULL, med_var = NULL, mle_avg_bias = NULL, med_avg_bias = NULL)

for (i in 1:length(dat_parameter$parameter)) {

 dat_sim_parameter <- rbind(dat_sim_parameter, estimators(1000, dat_parameter$parameter[i], 50))
}

dat_parameter <- cbind(dat_parameter, dat_sim_parameter)

mse_parameter <- dat_parameter %>%
  ggplot() +
  geom_path(aes(x = parameter, y = mle_mse), col = "red") +
  geom_path(aes(x = parameter, y = med_mse), col = "green") +
  geom_path(data = manyR50, aes(x = rate, y = mean_MSE), col = "yellow") +
  ylab("MSE") +
  xlab("Theta")

bias_parameter <- dat_parameter %>%
  ggplot() +
  geom_path(aes(x = parameter, y = mle_avg_bias), col = "red") +
  geom_path(aes(x = parameter, y = med_avg_bias), col = "green") +
  geom_path(data = manyR50, aes(x = rate, y = mean_bias), col = "yellow") +
  ylab("Bias") +
  xlab("Theta")

var_parameter <- dat_parameter %>%
  ggplot() +
  geom_path(aes(x = parameter, y = mle_var), col = "red") +
  geom_path(aes(x = parameter, y = med_var), col = "green") +
  geom_path(data = manyR50, aes(x = rate, y = mean_variance), col = "yellow") +
  ylab("Variance") +
  xlab("Theta")

grid.arrange(mse_parameter, bias_parameter, var_parameter, nrow = 3)

```


```{r}
dat_n <- data.frame(n_size = seq(5, 100, 1))

dat_sim_n <- data.frame(mle_mse = NULL, med_mse = NULL, mle_var = NULL, med_var = NULL, mle_avg_bias = NULL, med_avg_bias = NULL)

for (i in 1:length(dat_n$n_size)) {

 dat_sim_n <- rbind(dat_sim_n, estimators(1000, 5, dat_n$n_size[i]))
}

dat_n <- cbind(dat_n, dat_sim_n)

mse_n <- dat_n %>%
  ggplot() +
  geom_path(aes(x = n_size, y = mle_mse), col = "red") +
  geom_path(aes(x = n_size, y = med_mse), col = "green") +
  geom_path(data = manyN, aes(x = n, y = mean_MSE), col = "yellow") +
  ylab("MSE") +
  xlab("n-size")

bias_n <- dat_n %>%
  ggplot() +
  geom_path(aes(x = n_size, y = mle_avg_bias), col = "red") +
  geom_path(aes(x = n_size, y = med_avg_bias), col = "green") +
  geom_path(data = manyN, aes(x = n, y = mean_bias), col = "yellow") +
  ylab("Bias") +
  xlab("n-size")

var_n <- dat_n %>%
  ggplot() +
  geom_path(aes(x = n_size, y = mle_var), col = "red") +
  geom_path(aes(x = n_size, y = med_var), col = "green") +
  geom_path(data = manyN, aes(x = n, y = mean_variance), col = "yellow") +
  ylab("Variance") +
  xlab("n-size")

grid.arrange(mse_n, bias_n, var_n, nrow = 3)

```

